Carbonate chemistry at Ny-Ålesund

### 2021-03-22 awipev-CO2.Rmd
Reworked impacts of nuts

## Impact of nutrients
```{r impact of nutrients, echo=FALSE, warning=FALSE, message=FALSE}
# boxplot
ts_po4_bp <- d_all %>%
  ggplot(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = po4)) +
  #geom_violin(fill='lightblue', alpha=0.5, size=0.2, trim = TRUE, na.rm = TRUE) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3, na.rm = TRUE) +
  labs(y = "po4", x=NULL, title=NULL) +
  Mytheme() +
  theme(panel.grid.major.x = element_blank())

ts_Si_bp <- d_all %>%
  ggplot(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = si)) +
  #geom_violin(fill='lightblue', alpha=0.5, size=0.2, trim = TRUE, na.rm = TRUE) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3, na.rm = TRUE) +
  labs(y = "si", x=NULL, title=NULL) +
  Mytheme() +
  theme(panel.grid.major.x = element_blank())
ts_po4_bp + ts_Si_bp
# Effects of changes in po4 and si
# 1st cakculate mean values
means <- d_all %>% 
  group_by(montht) %>% 
  summarize(mean_ct = mean(ct, na.rm = TRUE),
            mean_at_calc = mean(at_calc, na.rm = TRUE),
            mean_pco2_calc = mean(pco2_calc_fb, na.rm = TRUE),
            mean_sal = mean(sal_fb, na.rm = TRUE),
            mean_temp = mean(temp_fb, na.rm = TRUE)) %>% 
  ungroup()
mean_ct <-  mean(means$mean_ct)
mean_at_calc <-  mean(means$mean_at_calc)
mean_pco2_calc <-  mean(means$mean_pco2_calc)
mean_sal <-  mean(means$mean_sal)
mean_temp <-  mean(means$mean_temp)
# then check impacts using pco2 and AT
carb_po4_1 <- carb(24, mean_pco2_calc, mean_at_calc*1e-6, S = mean_sal, T = mean_temp, P = 0, 
                 Pt = c(min(d_all$po4/1e6, na.rm=TRUE), max(d_all$po4/1e6, na.rm=TRUE)), Sit = 0,
                 k1k2 = "l", kf = "dg", ks = "d", pHscale = "T", b = "u74")
diff_po4_1 <- carb_po4_1[1,] - carb_po4_1[2,]
carb_Si_1 <- carb(24, mean_pco2_calc, mean_at_calc*1e-6, S = mean_sal, T = mean_temp, P = 0,
                Pt = 0, Sit = c(min(d_all$si/1e6, na.rm=TRUE), max(d_all$si/1e6, na.rm=TRUE)), 
                k1k2 = "l", kf = "dg", ks = "d", pHscale = "T", b = "u74")
diff_Si_1 <- carb_Si_1[1,] - carb_Si_1[2,]

# then check impacts using AT-DIC pair
carb_po4_2 <- carb(15, mean_at_calc*1e-6, mean_ct*1e-6, S = mean_sal, T = mean_temp, P = 0, 
                 Pt = c(min(d_all$po4/1e6, na.rm=TRUE), max(d_all$po4/1e6, na.rm=TRUE)), Sit = 0,
                 k1k2 = "l", kf = "dg", ks = "d", pHscale = "T", b = "u74")
diff_po4_2 <- carb_po4_2[1,] - carb_po4_2[2,]
carb_Si_2 <- carb(15, mean_at_calc*1e-6, mean_ct*1e-6, S = mean_sal, T = mean_temp, P = 0,
                Pt = 0, Sit = c(min(d_all$si/1e6, na.rm=TRUE), max(d_all$si/1e6, na.rm=TRUE)), 
                k1k2 = "l", kf = "dg", ks = "d", pHscale = "T", b = "u74")
diff_Si_2 <- carb_Si_2[1,] - carb_Si_2[2,]
rm(ts_po4_bp, ts_Si_bp)
```

@@

### Using the pCO2-AT pair (hourly data)
The impact of the min max changes of po4 on pH, DIC, and Omega_a are small:  `r round(1000*diff_po4_1$pH, digits=2)` mpH units, `r round(1e6*diff_po4_1$DIC, digits=2)` umol/kg for CT and `r round(diff_po4_1$OmegaAragonite, digits=2)` for OmegaAragonite.

The impact of the min max changes of si on pH, DIC, and Omega_a are small:  `r round(1000*diff_Si_1$pH, digits=2)` mpH units, `r round(1e6*diff_Si_1$DIC, digits=2)` umol/kg for CT and `r round(diff_Si_1$OmegaAragonite, digits=2)` for OmegaAragonite.

### Using the CT-AT pair (discrete data)
The impact of the min max changes of po4 on pH, pCO2, and Omega_a are small:  `r round(1000*diff_po4_2$pH, digits=2)` mpH units, `r round(diff_po4_2$pCO2, digits=2)` uatm for pCO2 and `r round(diff_po4_2$OmegaAragonite, digits=2)` for OmegaAragonite.

The impact of the min max changes of po4 on pH, pCO2, and Omega_a are small:  `r round(1000*diff_Si_2$pH, digits=2)` mpH units, `r round(diff_Si_2$pCO2, digits=2)` uatm for pCO2 and `r round(diff_Si_2$OmegaAragonite, digits=2)` for OmegaAragonite.

@@

### Lueker vs Sulpis using FerryBox pCO2 and at in situ
```{r impact of constants on FerryBox data, echo=FALSE, warning=FALSE, message=FALSE}
# Compare Lueker and Sulpis using pco2_fb and at in situ

@@

The discrete values of pCO2, salinity and temperature in the FerryBox and in situ AT were used to calculate pH from measured pCO2 and AT using the formulations for K1 and K2 of Lueker et al. (2000) and Sulpis et al. (2020). The differences induced by the choice of the carbonic acid constants is:

>>>>

## Impact of nutrients   {.tabset .tabset-pills}
```{r impact of nutrients, echo=FALSE, warning=FALSE, message=FALSE}
# boxplot
ts_po4_bp <- d_all %>%
  ggplot(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = po4)) +
  #geom_violin(fill='lightblue', alpha=0.5, size=0.2, trim = TRUE, na.rm = TRUE) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3, na.rm = TRUE) +
  labs(y = "po4", x=NULL, title=NULL) +
  Mytheme() +
  theme(panel.grid.major.x = element_blank())

ts_Si_bp <- d_all %>%
  ggplot(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = si)) +
  #geom_violin(fill='lightblue', alpha=0.5, size=0.2, trim = TRUE, na.rm = TRUE) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3, na.rm = TRUE) +
  labs(y = "si", x=NULL, title=NULL) +
  Mytheme() +
  theme(panel.grid.major.x = element_blank())
ts_po4_bp + ts_Si_bp
# Effects of changes in po4 and si
# 1st cakculate mean values
means <- d_all %>% 
  group_by(montht) %>% 
  summarize(mean_ct = mean(ct, na.rm = TRUE),
            mean_at_calc = mean(at_calc, na.rm = TRUE),
            mean_pco2_calc = mean(pco2_calc_fb, na.rm = TRUE),
            mean_sal = mean(sal_fb, na.rm = TRUE),
            mean_temp = mean(temp_fb, na.rm = TRUE)) %>% 
  ungroup()
mean_ct <-  mean(means$mean_ct)
mean_at_calc <-  mean(means$mean_at_calc)
mean_pco2_calc <-  mean(means$mean_pco2_calc)
mean_sal <-  mean(means$mean_sal)
mean_temp <-  mean(means$mean_temp)
# then check impacts of 0, min and max nutrient values using pco2 and AT
carb_po4_1 <- carb(24, mean_pco2_calc, mean_at_calc*1e-6, S = mean_sal, T = mean_temp, P = 0, 
                 Pt = c(0, min(d_all$po4/1e6, na.rm=TRUE), max(d_all$po4/1e6, na.rm=TRUE)), Sit = 0,
                 k1k2 = "l", kf = "dg", ks = "d", pHscale = "T", b = "u74")
diff_po4_0a <- carb_po4_1[1,] - carb_po4_1[3,] # diff between no PO4 and max PO4
diff_po4_1a <- carb_po4_1[2,] - carb_po4_1[3,] # diff between min PO4 and max PO4
carb_Si_1 <- carb(24, mean_pco2_calc, mean_at_calc*1e-6, S = mean_sal, T = mean_temp, P = 0,
                Pt = 0, Sit = c(0, min(d_all$si/1e6, na.rm=TRUE), max(d_all$si/1e6, na.rm=TRUE)), 
                k1k2 = "l", kf = "dg", ks = "d", pHscale = "T", b = "u74")
diff_Si_0a <- carb_Si_1[1,] - carb_Si_1[3,] # diff between no Si and max Si
diff_Si_1a <- carb_Si_1[2,] - carb_Si_1[3,] # diff between no Si and max Si

# then check impacts of min and max nutrient values using AT-DIC pair
carb_po4_2 <- carb(15, mean_at_calc*1e-6, mean_ct*1e-6, S = mean_sal, T = mean_temp, P = 0, 
                 Pt = c(0, min(d_all$po4/1e6, na.rm=TRUE), max(d_all$po4/1e6, na.rm=TRUE)), Sit = 0,
                 k1k2 = "l", kf = "dg", ks = "d", pHscale = "T", b = "u74")
diff_po4_0b <- carb_po4_2[1,] - carb_po4_2[3,] # diff between no PO4 and max PO4
diff_po4_1b <- carb_po4_2[2,] - carb_po4_2[3,] # diff between min PO4 and max PO4
carb_Si_2 <- carb(15, mean_at_calc*1e-6, mean_ct*1e-6, S = mean_sal, T = mean_temp, P = 0,
                Pt = 0, Sit = c(0, min(d_all$si/1e6, na.rm=TRUE), max(d_all$si/1e6, na.rm=TRUE)), 
                k1k2 = "l", kf = "dg", ks = "d", pHscale = "T", b = "u74")
diff_Si_0b <- carb_Si_2[1,] - carb_Si_2[3,] # diff between no Si and max Si
diff_Si_1b <- carb_Si_2[2,] - carb_Si_2[3,] # diff between no Si and max Si

#nuts <- tibble(Pair = "pCO2-AT", "CT-AT",
#)
nuts_table <- bind_rows(diff_po4_0a, diff_Si_0a, diff_po4_0b, diff_Si_0b) %>% 
  dplyr::mutate(flag = c("pCO2 and ALK", "pCO2 and ALK", "ALK and DIC", "ALK and DIC"),
                nutrients = c("No PO4-max PO4", "No Si-max Si", "No PO4-max PO4", 
                              "No Si-max Si")) %>% 
  dplyr::select(flag, nutrients, pH, pCO2, DIC, OmegaAragonite)
rm(ts_po4_bp, ts_Si_bp)
```
`r knitr::kable(nuts_table, digits = c(1, 1, 4, 2, 4, 3), caption = "Impact of PO4 and Si.", format = "simple")`

@@

### Using the pCO2-AT pair (hourly data)
The difference between no PO4 and its maximum value on pH, DIC, and Omega_a are small:  `r round(1000*diff_po4_0a$pH, digits=2)` mpH units, `r round(1e6*diff_po4_0a$DIC, digits=2)` umol/kg for CT and `r round(diff_po4_0a$OmegaAragonite, digits=2)` for OmegaAragonite.

The impact of the min max changes of po4 on pH, DIC, and Omega_a are small:  `r round(1000*diff_po4_1a$pH, digits=2)` mpH units, `r round(1e6*diff_po4_1a$DIC, digits=2)` umol/kg for CT and `r round(diff_po4_1a$OmegaAragonite, digits=2)` for OmegaAragonite.

The difference between no Si and its maximum value on pH, DIC, and Omega_a are small:  `r round(1000*diff_Si_0a$pH, digits=2)` mpH units, `r round(1e6*diff_Si_0a$DIC, digits=2)` umol/kg for CT and `r round(diff_Si_0a$OmegaAragonite, digits=2)` for OmegaAragonite.

The impact of the min max changes of si on pH, DIC, and Omega_a are small:  `r round(1000*diff_Si_1a$pH, digits=2)` mpH units, `r round(1e6*diff_Si_1a$DIC, digits=2)` umol/kg for CT and `r round(diff_Si_1a$OmegaAragonite, digits=2)` for OmegaAragonite.

### Using the CT-AT pair (discrete data)
The difference between no PO4 and its maximum value on pH, pCO2, and Omega_a are small:  `r round(1000*diff_po4_0b$pH, digits=2)` mpH units, `r round(diff_po4_0b$pCO2, digits=2)` uatm for pCO2 and `r round(diff_po4_0b$OmegaAragonite, digits=2)` for OmegaAragonite.

The impact of the min max changes of po4 on pH, pCO2, and Omega_a are small:  `r round(1000*diff_po4_1b$pH, digits=2)` mpH units, `r round(diff_po4_1b$pCO2, digits=2)` uatm for pCO2 and `r round(diff_po4_1b$OmegaAragonite, digits=2)` for OmegaAragonite.

The difference between no Si and its maximum value on pH, pCO2, and Omega_a are small:  `r round(1000*diff_Si_0b$pH, digits=2)` mpH units, `r round(diff_Si_0b$pCO2, digits=2)` uatm for pCO2 and `r round(diff_Si_0b$OmegaAragonite, digits=2)` for OmegaAragonite.

The impact of the min max changes of Si on pH, pCO2, and Omega_a are small:  `r round(1000*diff_Si_1b$pH, digits=2)` mpH units, `r round(diff_Si_1b$pCO2, digits=2)` uatm for pCO2 and `r round(diff_Si_1b$OmegaAragonite, digits=2)` for OmegaAragonite.

@@

### K1K2 using FerryBox pCO2 and at in situ
```{r impact of constants using FerryBox pCO2 and at in situ, echo=FALSE, warning=FALSE, message=FALSE}

@@

The discrete values of pCO2, salinity and temperature in the FerryBox and in situ AT were used to calculate pH from measured pCO2 and AT using the formulations for K1 and K2 of Lueker et al. (2000) and Sulpis et al. (2020). Note that this cannot be done with in situ AT and CT because spectrophotometric pH in situ is not at a fixed depth.

The differences induced by the choice of the carbonic acid constants is:
